---
title: "Tidy and Transform Data"
author: "Durley Torres-Marin"
date: "September 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
##install.packages("tidyverse")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)
##Data structure
arrival.delays <- read.csv("https://raw.githubusercontent.com/doradu8030/Data607/master/Arrival%20delays%20by%20Airline.csv", stringsAsFactor = FALSE)
arrival.delays
```
## Tyding the Data and Manipulating string
This chunk is replacing the character " " to "_" into the word "on time" 
Adding the airline name to the two rows whitout it
removing the empty row 
```{r}
arrival.delays[2,1] <- arrival.delays[1,1]
arrival.delays[5,1] <- arrival.delays[4,1]
arrival.delays[,2] <- sapply(arrival.delays[,2], str_replace, " ", "_")

## *removing the empty row*
arrival.delays <- na.omit(arrival.delays) 
arrival.delays
```
## Naming Columns
```{r}
colnames(arrival.delays)[1] <- "Airline"
colnames(arrival.delays)[2] <- "Status"
arrival.delays
```
## Melting the data : List of columns from 3 to 7 that fall into a variable "City"
```{r}
draft.table <- gather(arrival.delays, "City", "Count", 3:7)
draft.table$City <- str_replace(draft.table$City, "[.]", " ")
draft.table
```

#Analysis

Total flights delayed group by Airline and City 
```{r}
f.arrival.delayed <- spread(draft.table, Status, Count)
totaldelayed <- f.arrival.delayed %>%
  group_by(Airline, City) %>% 
  summarise(fligh.delay = sum(delayed))
totaldelayed
```
Total flights on-time group by Airline and City 
```{r}
f.arrival.ontime <- spread(draft.table, Status, Count)
totalontime <- f.arrival.delayed %>%
  group_by(Airline, City) %>% 
  summarise(flight.ontime = sum(on_time))
totalontime
```
Total of flights delayed and on-time by Airline & shows
What percentage of flights are delayed by Airline
What percentage of flights are on-time by Airline
```{r}
totalstatus <- f.arrival.delayed %>% 
  group_by(Airline) %>%
  summarise(tot_delayed=sum(delayed),tot_on_time=sum(on_time),Perc_delay=round((tot_delayed/(tot_delayed+tot_on_time))*100, 2),Perc_on_time=round((tot_on_time/(tot_delayed+tot_on_time))*100, 2))
totalstatus
```

