---
title: "Assigment1-Loading Data into a Data Frame"
author: "Durley Torres-Marin"
date: "August 31, 2017"
output: html_document
---
Data Set Information:
---
"This data set includes descriptions of hypothetical samples corresponding to 23 species of gilled mushrooms in the Agaricus and Lepiota Family (pp. 500-525). Each species is identified as definitely edible, definitely poisonous, or of unknown edibility and not recommended. This latter class was combined with the poisonous one"

This data set is as an example of Categorical data which assigns numbers or characters to represent different categories of answers. These types of data can't be arithmetically manipulated.

The best way to summarize categorical data is with frequencies.  Frequencies can also be translated into percentages to show what percentage of the total sample answered in each category. 

---
Importing Data as a CSV file - no header was found, Header=FALSE
---
```{r}
Mushrooms <- read.csv(url("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"), header = FALSE)
head.matrix(x = Mushrooms, 6L)
```
---
What type of thing data is
---
```{r}
class(Mushrooms)
```
```{r}
dim(Mushrooms)
```
---
Creating the data set with a subset with sequences
we can use the c() combine function to address non-sequential rows and columns.
---
```{r}
Mushrooms1 <- Mushrooms[,c(1,6,14,21,22,23)]
head.matrix(x = Mushrooms1, 6L)
```
---
The columns were selected base on the attributes and rules given to identify and or avoid poisonous mushrooms 

Attribute Information

classification : Poisonous or edible 
odor: almond=a, anise=l, creosote=c, fishy=y, foul=f, musty=m, none=n, pungent=p,           spicy=s
stalk-surface-below-ring: fibrous=f, scaly=y, silky=k, smooth=s
spore-print-color: black=k, brown=n, buff=b, chocolate=h, green=r, orange=o, purple=u,       white=w, yellow=y
population: abundant=a, clustered=c, numerous=n, scattered=s, several=v, solitary=y
Habitat : grasses=g,leaves=l, meadows=m, paths=p, urban=u, waste=w, woods=d
---
Renaming columns
```{r}
names(Mushrooms1) = c("classification", "odor", "stalk-sbr", "spore-pr-c", "population", "habitat")
names(Mushrooms1)
```
Subsetting data 
Data set of Edible Mushrooms -  4280 instances

```{r}
edible_mushrooms <- Mushrooms1
edible_mushrooms = subset(Mushrooms1, classification == "e")
head.matrix(x = edible_mushrooms, 6L)
```
Data set of Poisonous Mushrooms - 3750 instances

```{r}
poisonous_mushrooms <- Mushrooms1
poisonous_mushrooms = subset(Mushrooms1, classification == "p")
head.matrix(x = poisonous_mushrooms, 6L)
```
updating a subset value 'e' as 'edible

```{r}
edible_mushrooms$classification <- ifelse(edible_mushrooms$classification== 'e', 'edible', edible_mushrooms$classification)
head.matrix(x = edible_mushrooms, 3L)
```

The attribute that is the best predictive analysis might be habitat that can tell us where the poisonous mushrooms can be found mean while edible mushrooms. 

Frequency distribution analysis using Bar Graphs using X axis = classification and Y axis = percentage

could answer questions like;

How many edible or poisonous mushrooms are in the data set? 	Results in frecuencies and/or percentages
```{r}
count = table(Mushrooms1$classification)
t = as.data.frame(count)
names(t)[1] = 'classification'
t

```

What is the habitat with more frecuency for poisonous mushrooms or edible mushrooms?
Possible habitat values: grasses=g,leaves=l, meadows=m, paths=p, urban=u, waste=w, woods=d

```{r}
count = table(Mushrooms1$classification, Mushrooms1$habitat)
t = as.data.frame(count)
names(t)[1] = 'classification'
names(t)[2] = 'habitat'
t

```
What odor is more common in the poisonous mushrooms or edible mushrooms? 
Posible odor values: almond=a, anise=l, creosote=c, fishy=y, foul=f, musty=m, none=n, pungent=p, spicy=s
```{r}
count1 = table(Mushrooms1$classification, Mushrooms1$odor)
t = as.data.frame(count1)
names(t)[1] = 'classification'
names(t)[2] = 'odor'
t
```
After studying the data set, I found two relevant attributes, habitat and odor that can provide additional information to determine of whether a particular mushroom is poisonous or edible.
 