---
title: "ProjecT1-DATA607"
author: "Durley Torres-Marin"
date: "September 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(qdapTools)
library(stringi)
```
# Removing Special Characters in R from a text file
Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
Player's Name
Player's State
Total Number of Points
Player's Pre-Rating
and Average Pre Chess Rating of Opponents

For the first player, the information would be:Gary Hua, ON, 6.0,1794, 1605
1605 was calculated by using the pre-tournament opponents' ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

Below are different block of commands I try to use to create a table and manipulate special characters from the text file; however I found much easier work with a vectors. Those command also removes special characters from a text file. 

"tournament.res <- read.table("https://raw.githubusercontent.com/doradu8030/Data607/master/tournamentinfo.txt", header=FALSE, sep="\n")
tournament.res <- readLines("tournamentinfo.txt")
str_replace_all(tournament.res, "[[:punct:]]", " ")
tournament.res<- tournament.res[-(1:4)]
tournament.res<-setdiff(tournament.res,lines)
tournament.res<-str_replace_all(tournament.res," {2}","")
tournament.res<- str_replace_all(tournament.res, "([\n /])", "")
gsub("[^[:alnum:][:blank:]+?&/\\-]", "", tournament.res)
gsub("[[:punct:]]", " ", tournament.res)
regex ("[[:punct:]]", tournament.res)""


I use the command read.delim which removes most of the special characters from the text file

```{r}
tournament.res <- read.delim("https://raw.githubusercontent.com/doradu8030/Data607/master/tournamentinfo.txt", header=FALSE, sep="|", skip = 4, stringsAsFactors = FALSE)
head(tournament.res)
```
After several attempts I created only one data frame with all fields needed.
```{r}
tournament.rounds <- data.frame(
            playername = tournament.res$V2[c(TRUE, FALSE, FALSE)],
            playerstate = tournament.res$V1[c(FALSE, TRUE, FALSE)],
            totnumpoints = round(as.numeric(tournament.res$V3[c(TRUE, FALSE, FALSE)]), 1),
            round1 = as.numeric(str_extract(tournament.res$V4[c(TRUE, FALSE, FALSE)],regex("[:digit:]+"))), 
            round2 = as.numeric(str_extract(tournament.res$V5[c(TRUE, FALSE, FALSE)],regex("[:digit:]+"))),
            round3 = as.numeric(str_extract(tournament.res$V6[c(TRUE, FALSE, FALSE)],regex("[:digit:]+"))),
            round4 = as.numeric(str_extract(tournament.res$V7[c(TRUE, FALSE, FALSE)],regex("[:digit:]+"))),
            round5 = as.numeric(str_extract(tournament.res$V8[c(TRUE, FALSE, FALSE)],regex("[:digit:]+"))),
            round6 = as.numeric(str_extract(tournament.res$V9[c(TRUE, FALSE, FALSE)],regex("[:digit:]+"))),
            round7 = as.numeric(str_extract(tournament.res$V10[c(TRUE, FALSE, FALSE)],regex("[:digit:]+"))),
            prerating = integer(64),
            avgrateoppon = integer(64))
            # the definicion of interger only allows the value 64 which corresponde             to the number of rows in the data set
```
How the instruction c(TRUE,FALSE,FALSE) included above works; 
See the original file 
```{r}
head(tournament.res)
```
Note that the information needed to be extrated is found every third row; row 1,4,7 and so on. 
see the changes when I added logical values to the c vector
```{r}
playername = tournament.res$V2[c(TRUE)]
head(playername)
```
```{r}
playername = tournament.res$V2[c(TRUE, FALSE)]
head(playername)
```
```{r}
playername = tournament.res$V2[c(TRUE, FALSE, TRUE)]
head(playername)
```
Now I will convert as numeric the column "prerating"";  at firt extract the character Ex."R: 1553" and assig it to a variable then R allows youto convert this variable as numeric and extracted as numeric as well
                   
```{r}
test <- str_extract(tournament.res$V2[c(FALSE, TRUE, FALSE)], "R:[:blank:]+[:digit:]+")
tournament.rounds$prerating <- as.numeric(str_extract(test, "[:digit:]+"))
head(test)
head(tournament.rounds$prerating)
```
The final output file is;
```{r}
tournament.rounds
```
```{r}
for (i in 1:64) {
       tournament.rounds$round1[i] <- tournament.rounds$prerating[tournament.rounds$round1[i]]
       tournament.rounds$round2[i] <- tournament.rounds$prerating[tournament.rounds$round2[i]]
       tournament.rounds$round3[i] <- tournament.rounds$prerating[tournament.rounds$round3[i]]
       tournament.rounds$round4[i] <- tournament.rounds$prerating[tournament.rounds$round4[i]]
       tournament.rounds$round5[i] <- tournament.rounds$prerating[tournament.rounds$round5[i]]
       tournament.rounds$round6[i] <- tournament.rounds$prerating[tournament.rounds$round6[i]]
       tournament.rounds$round7[i] <- tournament.rounds$prerating[tournament.rounds$round7[i]]
}
```
```{r}
  for (i in 1:64) {
       tournament.rounds$avgrateoppon[i] <- rowSums(tournament.rounds[i, 4:10], na.rm = TRUE) %/% 7
  }

foutput <- tournament.rounds[c(1,2,3,11,12)]
names(foutput) = c("PlayerName","PlayerState","NumofPoints","Pre-Rating","AvgRatingOp")
foutput
```
Then after all information needed is cleaned up, the final file could be save as csv file that could be exported or copy into a postgreSQL or MySQL. 
Note the R allows to work with dots (.) or underscore (-) in between of a variable names Ex.player.state
However, postgreSQL will present this message error; 
       
       player.state in posgreSQL
       ERROR:  syntax error at or near "."
       LINE 5:   player.name character(20) NOT NULL,
                       
       
If the final file will be exported to postgreSQL don't use those separators, postgreSQL wouldn't recognized (.),(-) as part of a variable name.

```{r}
write.csv(foutput, file = "C:/Users/Dora/CUNY/DATA607-Data Acquisition/Project1/tournamentfinal.csv")
```
