---
title: "RandSQL"
author: "Durley Torres-Marin"
date: "September 8, 2017"
output: html_document
---
## How to import SQL information to a data frame, from postgreSQL 
PostgreSQL use in R RPostgrade package
Load DBI package: It is requiered
```{r}
library("DBI", lib.loc="~/R/win-library/3.4")
library("RPostgreSQL", lib.loc="~/R/win-library/3.4")
library(sqldf)
con = dbConnect(PostgreSQL(), user="postgres", password="Doradu8030",
                 host="localhost", port=5432, dbname="mysql")
```

I used two different methods to populate my tables in SQL

   First: I loaded a .csv file from postgreSQL using the instruction "COPY"
   script name : Assig2.sql
   
   Second: I used  the INSERT instruction  script name :Assigwinsert.sql

Being more eficient the first method especially for large files already populated. However, I found useful and faster the second method if a conditional and repetitive routine like IF or FOR will be added to populate tables more efficiently (automate process).  

For booth, the first and second method, I created two .sql scripts to show this. 

## Getting data from my first table 
```{r}
topmovies = dbGetQuery(con,"select * from movie_info ")
names(topmovies) = c("Movie_Title", "Date_Release"," Director")
topmovies
```
## Getting data from my second table 
```{r}
movierate = dbGetQuery(con,"select * from movie_rate ")
names(movierate) = c("Movie_Title", "Scale 1-5")
movierate
```
## I created a third table using INSERT INTO SELECT, it requires that data types in source and target tables match to show aditional information.

```{r}
result = dbGetQuery(con,"select * from movie_res ")
names(result) = c("Movie_Title", "Scale 1-5","Date_Release", "Director")
result
```
```{r}
X<-dbDisconnect(con)
X
```

