---
title: "XML and JSON"
author: "Durley Torres-Marin"
date: "October 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Working with XML and JSON inR 

###Load Packages
```{r}
knitr::opts_chunk$set(warning=FALSE, 
                      message=FALSE,
                      tidy=F,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))
library(XML)
library(rjson)
library(RCurl)
library(jsonlite)
library(RJSONIO)
suppressMessages(library(XML))
suppressWarnings(library(XML))
suppressMessages(library(rjson))
suppressMessages(library(jsonlite))
suppressWarnings(library(jsonlite))
suppressPackageStartupMessages(library(jsonlite))
suppressMessages(library(RJSONIO))
suppressWarnings(library(RJSONIO))
```


##Reading a HTML file
```{r}
books.html <- readLines("https://raw.githubusercontent.com/doradu8030/Data607/master/books.html")

###Finding the lines matching the pattern "<td>...</td>" 
html.pattern = "<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>"

###Tyding the file 
b.content = grep(html.pattern, books.html[1:length(books.html)], value = TRUE)
b.info <- trimws(unlist(sub("<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>", "\\1", b.content)))


```

__Convert in a Data.frame__
```{r}
b.infoDF <- cbind.data.frame(split(b.info, rep(1:5, times=length(b.info)/5)), stringsAsFactors=F)
names(b.infoDF) <- c("Title", "Author", "Topic", "Publication_Year", "ISBN")
b.infoDF = b.infoDF[-1,]

# HTML Contents as Dataframe
b.infoDF
```
##Reading a XML file
```{r}
url.xml = getURL("https://raw.githubusercontent.com/doradu8030/Data607/master/books2.XML")  

xml_df <- xmlToDataFrame(url.xml)
#book.xml<- xmlParse(xml_df)
xml_df
```
```{r}
download.file("https://raw.githubusercontent.com/doradu8030/Data607/master/books.XML", destfile = "book_XML.xml")
  
book_XML.xml <- xmlParse("book_XML.xml")
books.XMLDF <- xmlToDataFrame(book_XML.xml)
books.XMLDF
```
###Reading a JSON file

```{r eval = FALSE}
url.json <-"https://raw.githubusercontent.com/doradu8030/Data607/master/books3.json"
json.Df <- as.data.frame(fromJSON(url.json))

```

```{r eval = FALSE}
jsonData <- rjson::fromJSON(file="https://raw.githubusercontent.com/doradu8030/Data607/master/books3.json")
```

Conclusion

Afert tons of attemps HTML and XML files are identical and also they are less difficult to manipulate than Jason files. Eventhough, I was not able to show the comparision between them, I was able to see in R when i was running chunks individually that the JSON file sctructure is different that XML and HTML files. I left some of the commands to load and parse a JSON file beign that at firts those command were giving the expected result but i could not figure out why after few attemps those commands were given tons of errors. For sure, i learned how to get data from HTML, XML and JSON files and the difference between their structures. 

Also, I notice that the column's data type in JSON is 'chr'while in XML and HTML are 'factors.


