---
title: "Assignment Week 7"
author: "Durley Torres-Marin"
date: "October 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(XML)
library(rjson)
library(RCurl)
library(stringr)  
library(tidyr)
library(jsonlite)
library(RJSONIO)
library(prettydoc)
library(jsonlite)
```
###Load an HTML file
```{r}
url_html <- getURLContent("https://raw.githubusercontent.com/doradu8030/Data607/master/books.html")
writeLines(url_html)

html_df <- readHTMLTable(url_html, which = 1)
html_df
```

##Reading an HTML file
```{r}
books.html <- readLines("https://raw.githubusercontent.com/doradu8030/Data607/master/books.html")

td.pattern = "<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>"

b.content = grep(td.pattern, books.html[1:length(books.html)], value = TRUE)

b.info <- trimws(unlist(sub("<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>", "\\1", b.content)))
b.info
```
##Structure of HTML
```{r}
str(books.html)
```
```{r}
b.infoDF <- cbind.data.frame(split(b.info, rep(1:5, times=length(b.info)/5)), stringsAsFactors=F)
names(b.infoDF) <- c("Title", "Author", "Theme", "Publication_Year", "ISBN")
b.infoDF = b.infoDF[-1,]

# HTML Contents as Dataframe
b.infoDF
```
```{r}
str(b.infoDF)
```
##Reading an XML file
```{r}
#books.json = JSON.parse()
download.file("https://raw.githubusercontent.com/doradu8030/Data607/master/books.XML", destfile = "book_XML.xml")
  
book_XML.xml <- xmlParse("book_XML.xml")
books.XMLDF <- xmlToDataFrame(book_XML.xml)
```
```{r}
str(books.XMLDF)
```
##Reading an JSON file
```{r}
books.json <- getURLContent("https://raw.githubusercontent.com/doradu8030/Data607/master/books1.json")
writeLines(books.json)
```
Conclusion

Afert tons of attemps HTML and XML files are identical and also they are less difficult to manipulate than Jason files. Eventhough, I was not able to show the comparision between them, I was able to see in R when i was running chunks individually that the JSON file sctructure is different that XML and HTML files. I left some of the commands to load and parse a JSON file beign that at firts those command were giving the expected result but i could not figure out why after few attemps those commands were given tons of errors. For sure, i learned how to get data from HTML, XML and JSON files and the difference between their structures. 

Also, I notice that the column's data type in JSON is 'chr'while in XML and HTML are 'factors.
```{r eval = FALSE}
json.Df <- as.data.frame(fromJSON(books.json))
```
